##General Info: \# - Author Name: Abigail Rangel \# - Data: CMPS 2016 \# - Project: RBSI 2023

##Packages #I will be installing packages to help with my dataset

```{r}
library(dplyr)
library(tidyr)
library(haven)
library(stargazer)
library(ggplot2)
library(estimatr)
library(ggeffects)
```

## Formatting My Code

####I will be loading my data into under a new name of "cmps_data"

```{r}
cmps_data <- da38040.0001
```

####Creating my new varaible of people who are on parole or probation

```{r}
cmps_data <- cmps_data %>% mutate(incar_ppl = ifelse(C289== "(1) Yes, I have", 1,
ifelse(C289== "(2) Yes, someone I know", 0 ,0)))
incar_data <- subset(cmps_data, incar_ppl == 1)

##Remember to add race such as latine and Black back into this data set
```

####Adding levels of disenfranchisment to my incar_data set to see if state has any impact
```{r}

# State1 is states that have no restrictions on voting, and states that have restrictions but only if your currently incarcerated (24 states + DC)
incar_data <- incar_data %>% mutate(state1 = ifelse(S4== "ME" | S4 == "VT"  | S4 == "DC"
| S4=="CA" |S4=="MI" | S4=="CO" | S4=="CT" | S4=="HI" | S4=="IL" | S4=="IN" | S4=="MD" | S4=="MA" | S4=="MT" | S4=="NV" | S4=="NH" | S4=="NJ" | S4=="NY" | S4=="NC" | S4=="ND" | S4=="OH" | S4=="OR" | S4=="PA" | S4=="RI" | S4=="UT" | S4=="WA", 1,
ifelse(S4=="AK" | S4=="AR"  | S4=="GA"  | S4=="ID"  | S4=="KS"  | S4=="LA"  | S4=="MN"  | S4=="MO"  | S4=="NM"  | S4=="OK"  | S4=="SC" | S4=="SD"  | S4=="TX"  | S4=="WV"  | S4=="WI" | S4=="AL" | S4=="AZ" | S4=="DE" | S4=="FL" | S4=="IA" | S4=="KY" | S4=="MS" | S4=="NE" | S4=="TN" | S4=="VA" | S4=="WY", 0, 0)))

# State2 is states that deny voting rights to people on/in prison, parole, and probation (15 states)
incar_data <- incar_data %>% mutate(state2 = ifelse(S4== "ME" | S4 == "VT"  | S4 == "DC"
| S4=="CA" |S4=="MI" | S4=="CO" | S4=="CT" | S4=="HI" | S4=="IL" | S4=="IN" | S4=="MD" | S4=="MA" | S4=="MT" | S4=="NV" | S4=="NH" | S4=="NJ" | S4=="NY" | S4=="NC" | S4=="ND" | S4=="OH" | S4=="OR" | S4=="PA" | S4=="RI" | S4=="UT" | S4=="WA", 0,
ifelse(S4=="AK" | S4=="AR"  | S4=="GA"  | S4=="ID"  | S4=="KS"  | S4=="LA"  | S4=="MN"  | S4=="MO"  | S4=="NM"  | S4=="OK"  | S4=="SC" | S4=="SD"  | S4=="TX"  | S4=="WV"  | S4=="WI" , 1,
ifelse(S4=="AL" | S4=="AZ" | S4=="DE" | S4=="FL" | S4=="IA" | S4=="KY" | S4=="MS" | S4=="NE" | S4=="TN" | S4=="VA" | S4=="WY", 0, 0))))

# State3 is states that deny voting rights to people on/in prison, parole, probation, and post-sentence (11 states)
incar_data <- incar_data %>% mutate(state3 = ifelse(S4== "ME" | S4 == "VT"  | S4 == "DC"
| S4=="CA" |S4=="MI" | S4=="CO" | S4=="CT" | S4=="HI" | S4=="IL" | S4=="IN" | S4=="MD" | S4=="MA" | S4=="MT" | S4=="NV" | S4=="NH" | S4=="NJ" | S4=="NY" | S4=="NC" | S4=="ND" | S4=="OH" | S4=="OR" | S4=="PA" | S4=="RI" | S4=="UT" | S4=="WA", 0,
ifelse(S4=="AK" | S4=="AR"  | S4=="GA"  | S4=="ID"  | S4=="KS"  | S4=="LA"  | S4=="MN"  | S4=="MO"  | S4=="NM"  | S4=="OK"  | S4=="SC" | S4=="SD"  | S4=="TX"  | S4=="WV"  | S4=="WI" , 0,
ifelse(S4=="AL" | S4=="AZ" | S4=="DE" | S4=="FL" | S4=="IA" | S4=="KY" | S4=="MS" | S4=="NE" | S4=="TN" | S4=="VA" | S4=="WY", 1, 0))))


```


####Adding dummies for my cmps_data for race


```{r}
##Controls
####Adding dummies for my controls in the cmps_data and incar_data

#Gender
#   - 0, Male
#   - 1, Female
cmps_data <- cmps_data %>% mutate(gender = ifelse(S3== "(1) Female", 1,
ifelse(S3== "(2) Male", 0 ,0)))

incar_data <- incar_data %>% mutate(gender = ifelse(S3== "(1) Female", 1,
ifelse(S3== "(2) Male", 0 ,0)))

#I am not creating dummy variables us as control for my data
#     - 0, not registered
#     - 1, registered
cmps_data <- cmps_data %>% mutate(reg_vote = ifelse(S5== "(1) Yes", 1,
ifelse(S5== "(2) No, not registered", 0 ,0)))

incar_data <- incar_data %>% mutate(reg_vote = ifelse(S5== "(1) Yes", 1,
ifelse(S5== "(2) No, not registered", 0 ,0)))

# Party identification
#     - 1, Rep
#     - 2, Dem 
#     - 3, Ind
#     - 4, Other
cmps_data <- cmps_data %>% mutate(party_id = ifelse(C25== "(1) Republican", 1,
ifelse(C25== "(2) Democrat", 2,
ifelse(C25== "(3) Independent", 3,
ifelse(C25== "(4) Other party", 4,0)))))

incar_data <- incar_data %>% mutate(party_id = ifelse(C25== "(1) Republican", 1,
ifelse(C25== "(2) Democrat", 2,
ifelse(C25== "(3) Independent", 3,
ifelse(C25== "(4) Other party", 4,0)))))

#Race
cmps_data <- cmps_data %>% mutate(race = ifelse(ETHNIC_QUOTA== "(1) White, Not-Hispanic", 1,
ifelse(ETHNIC_QUOTA== "(2) Hispanic or Latino", 2,
ifelse(ETHNIC_QUOTA== "(3) Black or African American", 3,
ifelse(ETHNIC_QUOTA== "(4) Asian American", 4,
ifelse(ETHNIC_QUOTA== "(5) Middle Eastern or Arab", 5,
ifelse(ETHNIC_QUOTA== "(6) American Indian/Native American", 6, 0)))))))

incar_data <- incar_data %>% mutate(race = ifelse(ETHNIC_QUOTA== "(1) White, Not-Hispanic", 1,
ifelse(ETHNIC_QUOTA== "(2) Hispanic or Latino", 2,
ifelse(ETHNIC_QUOTA== "(3) Black or African American", 3,
ifelse(ETHNIC_QUOTA== "(4) Asian American", 4,
ifelse(ETHNIC_QUOTA== "(5) Middle Eastern or Arab", 5,
ifelse(ETHNIC_QUOTA== "(6) American Indian/Native American", 6, 0)))))))

#Races
------------------------
#Latine
incar_data <- incar_data %>% mutate(latine = ifelse(ETHNIC_QUOTA== "(2) Hispanic or Latino", 1,0))

#Black
incar_data <- incar_data %>% mutate(Black = ifelse(ETHNIC_QUOTA== "(3) Black or African American", 1,0))

#POC 
incar_data <- incar_data %>% mutate(POC = ifelse(ETHNIC_QUOTA== "(3) Black or African American"|ETHNIC_QUOTA=="(2) Hispanic or Latino"|ETHNIC_QUOTA=="(4) Asian American"|ETHNIC_QUOTA=="(5) Middle Eastern or Arab"|ETHNIC_QUOTA=="(6) American Indian/Native American", 1,0))


# Education
#     - 1, Grades 1 - 8
#     - 2, Some high school
#     - 3, High school graduate or GED
#     - 4, Some college, 2-year degree
#     - 5, 4-year college graduate
#     - 6, Post-graduate education
cmps_data <- cmps_data %>% mutate(educ = ifelse(C381== "(1) Grades 1 - 8", 1,
ifelse(C381== "(2) Some high school", 2,
ifelse(C381== "(3) High school graduate or GED", 3,
ifelse(C381== "(4) Some college, 2-year degree", 4,
ifelse(C381== "(5) 4-year college graduate", 5,
ifelse(C381== "(6) Post-graduate education", 6,0)))))))

incar_data <- incar_data %>% mutate(educ = ifelse(C381== "(1) Grades 1 - 8", 1,
ifelse(C381== "(2) Some high school", 2,
ifelse(C381== "(3) High school graduate or GED", 3,
ifelse(C381== "(4) Some college, 2-year degree", 4,
ifelse(C381== "(5) 4-year college graduate", 5,
ifelse(C381== "(6) Post-graduate education", 6,0)))))))

# Education
# scale 1 being the lowest and 12 the highest
cmps_data <- cmps_data %>% mutate(income = ifelse(C383== "(01) Less than $20,000", 1,
ifelse(C383== "(02) $20,000 to $29,999", 2,
ifelse(C383== "(03) $30,000 to $39,999", 3,
ifelse(C383== "(04) $40,000 to $49,999", 4,
ifelse(C383== "(05) $50,000 to $59,999", 5,
ifelse(C383== "(06) $60,000 to $69,999", 6,
ifelse(C383== "(07) $70,000 to $79,999", 7,
ifelse(C383== "(08) $80,000 to $89,999", 8,
ifelse(C383== "(09) $90,000 to $99,999", 9,
ifelse(C383== "(10) $100,000 to $149,999", 10,
ifelse(C383== "(11) $150,000 to $199,999", 11,
ifelse(C383== "(12) $200,000 or more", 12,NA)))))))))))))

incar_data <- incar_data %>% mutate(income = ifelse(C383== "(01) Less than $20,000", 1,
ifelse(C383== "(02) $20,000 to $29,999", 2,
ifelse(C383== "(03) $30,000 to $39,999", 3,
ifelse(C383== "(04) $40,000 to $49,999", 4,
ifelse(C383== "(05) $50,000 to $59,999", 5,
ifelse(C383== "(06) $60,000 to $69,999", 6,
ifelse(C383== "(07) $70,000 to $79,999", 7,
ifelse(C383== "(08) $80,000 to $89,999", 8,
ifelse(C383== "(09) $90,000 to $99,999", 9,
ifelse(C383== "(10) $100,000 to $149,999", 10,
ifelse(C383== "(11) $150,000 to $199,999", 11,
ifelse(C383== "(12) $200,000 or more", 12,NA)))))))))))))
```

##Indepedent Varaibles

```{r}
## Now I am going to fix the code for my dependent variables, Most of theses are on scales

#c47. How much of the time do you trust the Federal government in Washington D.C. to do what is right?
cmps_data <- cmps_data %>% mutate(gov_cares = ifelse(C47== "(1) Just about always", 4,
ifelse(C47== "(2) Most of the time", 3,
ifelse(C47== "(3) Only sometimes", 2,
ifelse(C47== "(4) Never at all", 1,0)))))

incar_data <- incar_data %>% mutate(gov_cares = ifelse(C47== "(1) Just about always", 4,
ifelse(C47== "(2) Most of the time", 3,
ifelse(C47== "(3) Only sometimes", 2,
ifelse(C47== "(4) Never at all", 1,0)))))

#C31 When it comes to politics, do you think of yourself as liberal, moderate, conservative
cmps_data <- cmps_data %>% mutate(self_id = ifelse(C31== "(1) Very Liberal", 1,
ifelse(C31== "(2) Somewhat Liberal", 2,
ifelse(C31== "(3) Moderate", 3,
ifelse(C31== "(4) Somewhat Conservative", 4,
ifelse(C31== "(5) Very Conservative", 5 ,0))))))

incar_data <- incar_data %>% mutate(self_id = ifelse(C31== "(1) Very Liberal", 1,
ifelse(C31== "(2) Somewhat Liberal", 2,
ifelse(C31== "(3) Moderate", 3,
ifelse(C31== "(4) Somewhat Conservative", 4,
ifelse(C31== "(5) Very Conservative", 5 ,0))))))

#bl171. How important do you think it is to vote, compared to other political activities? 
#ONLY ASKED TO LATINO AND BLACK
cmps_data <- cmps_data %>% mutate(voting_import = ifelse(BL171== "(1) The most important", 5,
ifelse(BL171== "(2) More important than most", 4,
ifelse(BL171== "(3) As important as other activities", 3,
ifelse(BL171== "(4) Not as important", 2,
ifelse(BL171== "(5) Not important at all", 1 ,0))))))

incar_data <- incar_data %>% mutate(voting_import = ifelse(BL171== "(1) The most important", 5,
ifelse(BL171== "(2) More important than most", 4,
ifelse(BL171== "(3) As important as other activities", 3,
ifelse(BL171== "(4) Not as important", 2,
ifelse(BL171== "(5) Not important at all", 1 ,0))))))

#C53 Do you participate in one or more than one social, cultural, civic, political group or union or do you not participate in the activities of any such groups?
cmps_data <- cmps_data %>% mutate(civil_part = ifelse(C53== "(2) Yes, more than one", 3,
ifelse(C53== "(1) Yes, one", 2,
ifelse(C53== "(3) None", 1 ,0))))

incar_data <- incar_data %>% mutate(civil_part = ifelse(C53== "(2) Yes, more than one", 3,
ifelse(C53== "(1) Yes, one", 2,
ifelse(C53== "(3) None", 1 ,0))))

#C66 In the last twelve months, have you: Attended a protest march, demonstration, or rally?
cmps_data <- cmps_data %>% mutate(attend_protest = ifelse(C66== "(1) Yes", 1,
ifelse(C66== "(2) No", 0, NA)))

incar_data <- incar_data %>% mutate(attend_protest = ifelse(C66== "(1) Yes", 1,
ifelse(C66== "(2) No", 0, NA)))

#C106 How much do you agree or disagree with the following statement: public officials don’t care much what people like me think?
cmps_data <- cmps_data %>% mutate(politician_cares = ifelse(C106== "(1) Strongly agree", 5,
ifelse(C106== "(2) Agree", 4,
ifelse(C106== "(3) Neither agree nor disagree", 3,
ifelse(C106== "(4) Disagree", 2,
ifelse(C106== "(5) Strongly disagree", 1 ,0))))))

incar_data <- incar_data %>% mutate(politician_cares = ifelse(C106== "(1) Strongly agree", 5,
ifelse(C106== "(2) Agree", 4,
ifelse(C106== "(3) Neither agree nor disagree", 3,
ifelse(C106== "(4) Disagree", 2,
ifelse(C106== "(5) Strongly disagree", 1 ,0))))))

#C233 How effective, if at all, are the following tactics for getting your voice heard by elected officials: Voting
cmps_data <- cmps_data %>% mutate(voting_effect = ifelse(C233== "(1) Very effective", 4,
ifelse(C233== "(2) Somewhat effective", 3,
ifelse(C233== "(3) Not too effective", 2,
ifelse(C233== "(4) Not at all effective", 1,NA)))))

incar_data <- incar_data %>% mutate(voting_effect = ifelse(C233== "(1) Very effective", 4,
ifelse(C233== "(2) Somewhat effective", 3,
ifelse(C233== "(3) Not too effective", 2,
ifelse(C233== "(4) Not at all effective", 1,NA)))))

#C234 How effective, if at all, are the following tactics for getting your voice heard by elected officials: Non-violent protest
cmps_data <- cmps_data %>% mutate(protest_effect = ifelse(C234== "(1) Very effective", 4,
ifelse(C234== "(2) Somewhat effective", 3,
ifelse(C234== "(3) Not too effective", 2,
ifelse(C234== "(4) Not at all effective", 1,NA)))))

incar_data <- incar_data %>% mutate(protest_effect = ifelse(C234== "(1) Very effective", 4,
ifelse(C234== "(2) Somewhat effective", 3,
ifelse(C234== "(3) Not too effective", 2,
ifelse(C234== "(4) Not at all effective", 1,NA)))))

#C236 How effective, if at all, are the following tactics for getting your voice heard by elected officials: riot
cmps_data <- cmps_data %>% mutate(riot_effect = ifelse(C236== "(1) Very effective", 4,
ifelse(C236== "(2) Somewhat effective", 3,
ifelse(C236== "(3) Not too effective", 2,
ifelse(C236== "(4) Not at all effective", 1,NA)))))

incar_data <- incar_data %>% mutate(riot_effect = ifelse(C236== "(1) Very effective", 4,
ifelse(C236== "(2) Somewhat effective", 3,
ifelse(C236== "(3) Not too effective", 2,
ifelse(C236== "(4) Not at all effective", 1,NA)))))
```

#Regressions/Analysis
```{r}
# Seeing if government cares about people like you across incarcerated and then within incarcerated people
m2 <- lm(formula = gov_cares ~ state2 + state3  + gender + AGE + income + party_id + educ, data = incar_data)
stargazer(m2, type = "text")
```

```{r}
#Seeing who is most likely to see if incarceration impacts your ideology
si2 <- lm(formula = self_id ~ state2 + state3 + gender + AGE + income + party_id + educ, data = incar_data)
stargazer(si2, type = "text")
```

```{r}
#Seeing who is most likely to see if incarceration impacts how important voting is
vi2 <- lm(formula = voting_import ~ state2 + state3 + gender + AGE + income + party_id + educ, data = incar_data)
stargazer(vi2, type = "text")
```

```{r}
#Seeing who is most likely to see if incarceration impacts if you participated in a civic
cp2 <- lm(formula = civil_part ~ state2 + state3 + gender + AGE + income + party_id + educ, data = incar_data)
stargazer(cp2, type = "text")
```

```{r}
#Seeing who is attends a protest, you are more likely to attend if you have been incarcerated 
ap2 <- lm(formula = attend_protest ~ state2 + state3 + gender + AGE + income + party_id + educ, data = incar_data)
stargazer(ap2, type = "text")
```

```{r}
#if polticans care about
pc2 <- lm(formula = politician_cares ~ state2 + state3 + gender + AGE + income + party_id + educ, data = incar_data)
stargazer(pc2, type = "text")
```

```{r}
#seeing how much voting has an effect
ve2 <- lm(formula = voting_effect ~ state2 + state3 + gender + AGE + income + party_id + educ + POC, data = incar_data)
stargazer(ve2, type = "text")
```

```{r}
pe2 <- lm(formula = protest_effect ~ state2 + state3 + gender + AGE + income + party_id + educ + POC, data = incar_data)
stargazer(pe2, type = "text")
```

```{r}
re2 <- lm(formula = riot_effect ~ state2 + state3  + gender + AGE + income + party_id + educ + POC, data = incar_data)
stargazer(re2, type = "text")
```
##Visuals - Tables/Graphs
```{r}
# Comparison of my 1st Hypothesis
stargazer(ve2, pe2, re2, covariate.labels = c("Disenfranchisment 2", "Disenfranchisment 3", "Gender", "Age", "Income", "Party ID", "Education", "POC"), dep.var.labels = c("Voting", "Protest", "Riot"), type = "html", out = "voting_protest_riot_compared.html", digits = 2)
```
```{r}
#Effectiveness of methods to get elected officials to recognize you
stargazer(ve2, pe2, re2, covariate.labels = c("Disenfranchisment 2", "Disenfranchisment 3", "Gender", "Age", "Income", "Party ID", "Education"), dep.var.labels = c("Voting", "Protest", "Riot"), type = "html", out = "effective_methods_states_lvl.html", digits = 2)
```
```{r}
#Tables for data regarding government caring
effects <- ggpredict(m2,                                            # model
             terms = c("state2", "state3"))                   # variables you want to graph
p5 <- ggplot(effects, aes(x = x, y = predicted, color = group)) + 
   geom_line() +
   scale_color_manual(labels = c("Mid Disenfranchisment States", "Worst Disenfranchisment States"), values = c("purple", "red")) +
      labs(x = "Incarcerated People",
        y = "Government Caring",
        title = "Correlation Between How Much Government Cares About 'People Like Me'", 
        caption = "Source: CMPS Data, 2016") +
   theme_classic()
p5
```

```{r}
effects <- ggpredict(vi2,                                            # model
             terms = c("state2", "state3"))                   # variables you want to graph
p2 <- ggplot(effects, aes(x = x, y = predicted, color = group)) + 
   geom_line() +
   scale_color_manual(labels = c("Mid Disenfranchisment States", "Worst Disenfranchisment States"), values = c("purple", "red")) +
      labs(x = "Incarcerated People",
        y = "Importance of Voting",
        title = "Correlation Between Importance of Voting Levels of State Disenfranchisment", 
        caption = "Source: CMPS Data, 2016") +
   theme_classic()
p2
```

```{r}
out <- summary(si2) # Create a placeholder for the summary results
out$coefficients

n <- nrow(out$coefficients)                                                           # Number of coefficients in the model
# Create a placeholder matrix, make sure you have 
coefMat <- matrix(NA, n-1, 4)
coefMat[,1] <- out$coefficients[2:n,1]
coefMat[,2] <- out$coefficients[2:n,2]
coefMat[,3] <- coefMat[,1] - qnorm(1-0.05/2)*coefMat[,2]
coefMat[,4] <- coefMat[,1] + qnorm(1-0.05/2)*coefMat[,2]
rownames(coefMat) <- c("Mid Disenfrancimsent States", "Worst Disenfrancimsent States", "Race", "Gender", "Income", "AGE","Party Identification", "Education")
colnames(coefMat) <- c("coef", "std_err", "min", "max")
df <- as.data.frame(coefMat)
# Plot
coefplot <- ggplot(data = df, aes(x = rownames(coefMat),
                           y = coef)) + 
   geom_hline(yintercept=0, colour="black", linewidth=1, linetype="dotted") +
   geom_linerange(aes(ymin=min, ymax=max), linewidth=0.2, color="grey") + 
   geom_point(color="blue") + 
   labs(x="Coefficient", y="Ideological Scale", title="Ideological Identificaiton of Formely Incarcerated Groups by Degree of Disenfrancimsent") +  
   coord_flip() + theme_classic()

coefplot
```

```{r}
stargazer(m2, pc2, covariate.labels = c("Disenfranchisment 2", "Disenfranchisment 3", "Gender", "Age", "Income", "Party ID", "Education"), dep.var.labels = c("Government Cares", "Politican Cares"), type = "html", out = "disillusionment.html", digits = 2)
```
```{r}
stargazer(ap2, cp2, covariate.labels = c("Disenfranchisment 2", "Disenfranchisment 3", "Gender", "Age", "Income", "Party ID", "Education"), dep.var.labels = c("Attended Protest", "Civil Participation" ), type = "html", out = "part_outcomes.html", digits = 2)
```

